\bf Approximating Likelihood Ratios with Calibrated Discriminative Classifiers

### 2016-03-17 jasa/paper-jasa.tex
fix figure refs

Finally, we evaluate the diagnostics described in Sec.~\ref{S:diagnostics} for this example. To aid in visualization, we restrict to a 1-dimensional slice of the likelihood along $\alpha$ with $\beta=1$.   We consider three situations: i) a poorly trained, but well calibrated classifier; ii) a well trained, but poorly calibrated classifier; and iii) a well trained, and well calibrated classifier. For each case, we employ two diagnostic tests. The first checks for independence of $-2\log\Lambda(\theta)$ with respect to changes in the reference value $\theta_1$ as shown in Eqn.~\ref{eq:theta_1_independence}. The second uses a classifier to distinguish between samples from  $p(\mathbf{x}|\theta_0)$ and samples from $p(\mathbf{x}|\theta_1)$ weighted according to $r(\mathbf{x}; \theta_0, \theta_1)$.  As discussed for Fig.~\ref{fig:5b}, statistical fluctuations in the calibration lead to some noise in the raw approximate likelihood. Thus, we show the posterior mean of a Gaussian processes resulting from Bayesian optimization of  the raw approximate likelihood as in Fig.~\ref{fig:5c}. In addition, the standard deviation of the Gaussian process is shown for one of the $\theta_1$ reference points to indicate the size of these statistical fluctuations. It is clear that in the well calibrated cases that these fluctuations are small, while in the poorly calibrated case these fluctuations are large. Moreover, in Fig.~\ref{fig:5a} we see that in the poorly trained, well calibrated case the classifier $\hat{s}(\mathbf{x}; \theta_0, \theta_1)$ has a significant dependence on the $\theta_1$ reference point. In contrast, in Fig.~\ref{fig:5c} the likelihood curves vary significantly, but this is comparable to the fluctuations expected from the calibration procedure. Finally, Fig.~\ref{fig:5e} shows that in the well trained, well calibrated case that the likelihood curves are all consistent with the exact likelihood within the estimated uncertainty band of the Gaussian process.  The ROC curves tell a similarly revealing story. As expected, the classifier is not able to distinguish between the distributions when $p(\mathbf{x}|\theta_1)$ is weighted by the exact likelihood ratio. We can also rule out that this is a deficiency in the classifier because the two distributions are well separated when no weights are applied to $p(\mathbf{x}|\theta_1)$. In both Fig.~\ref{fig:5b} and Fig.~\ref{fig:5d} the ROC curve correctly diagnoses deficiencies in the approximate likelihood ratio  $\hat{r}(\hat{s}(\mathbf{x}; \theta_0, \theta_1))$. Finally, Fig.~\ref{fig:5f} shows that the ROC curve in the well trained, well calibrated case is almost identical with the exact likelihood ratio, confirming the quality of the approximation.

@@

%\subsubsection{Diagnostics}

%\begin{figure}
%    \centering
%    \includegraphics[height=.9\textheight]{figures/diagnostics.pdf}
%    \caption{Diagnostic plots ....}
%    \label{fig:ratio_comp}
%\end{figure}

>>>>

Finally, we evaluate the diagnostics described in Sec.~\ref{S:diagnostics} for this example. To aid in visualization, we restrict to a 1-dimensional slice of the likelihood along $\alpha$ with $\beta=1$.   We consider three situations: i) a poorly trained, but well calibrated classifier; ii) a well trained, but poorly calibrated classifier; and iii) a well trained, and well calibrated classifier. For each case, we employ two diagnostic tests. The first checks for independence of $-2\log\Lambda(\theta)$ with respect to changes in the reference value $\theta_1$ as shown in Eqn.~\ref{eq:theta_1_independence}. The second uses a classifier to distinguish between samples from  $p(\mathbf{x}|\theta_0)$ and samples from $p(\mathbf{x}|\theta_1)$ weighted according to $r(\mathbf{x}; \theta_0, \theta_1)$.  As discussed for Fig.~\ref{fig:5b}, statistical fluctuations in the calibration lead to some noise in the raw approximate likelihood. Thus, we show the posterior mean of a Gaussian processes resulting from Bayesian optimization of  the raw approximate likelihood as in Fig.~\ref{fig:6c}. In addition, the standard deviation of the Gaussian process is shown for one of the $\theta_1$ reference points to indicate the size of these statistical fluctuations. It is clear that in the well calibrated cases that these fluctuations are small, while in the poorly calibrated case these fluctuations are large. Moreover, in Fig.~\ref{fig:6a} we see that in the poorly trained, well calibrated case the classifier $\hat{s}(\mathbf{x}; \theta_0, \theta_1)$ has a significant dependence on the $\theta_1$ reference point. In contrast, in Fig.~\ref{fig:6c} the likelihood curves vary significantly, but this is comparable to the fluctuations expected from the calibration procedure. Finally, Fig.~\ref{fig:6e} shows that in the well trained, well calibrated case that the likelihood curves are all consistent with the exact likelihood within the estimated uncertainty band of the Gaussian process.  The ROC curves tell a similarly revealing story. As expected, the classifier is not able to distinguish between the distributions when $p(\mathbf{x}|\theta_1)$ is weighted by the exact likelihood ratio. We can also rule out that this is a deficiency in the classifier because the two distributions are well separated when no weights are applied to $p(\mathbf{x}|\theta_1)$. In both Fig.~\ref{fig:5b} and Fig.~\ref{fig:6d} the ROC curve correctly diagnoses deficiencies in the approximate likelihood ratio  $\hat{r}(\hat{s}(\mathbf{x}; \theta_0, \theta_1))$. Finally, Fig.~\ref{fig:6f} shows that the ROC curve in the well trained, well calibrated case is almost identical with the exact likelihood ratio, confirming the quality of the approximation.

@@
